### 2018-04-02 lib/coverage.Rmd
progress

---
title: "Open Policing Project - City Coverage"
classoption: landscape
---

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(kableExtra)
source("coverage.R")
cvg <- coverage() %>% select(-path, -modified_time, -state, -incident_id)
kable(cvg, row.names = TRUE, format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = "scale_down")
```