### 2018-08-21 lib/coverage.Rmd
progress

---
title: "Open Policing Project - City: Coverage Rates"
classoption: landscape
---

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(kableExtra)
source("coverage.R")
cvg <- coverage() %>%
  select(-path, -modified_time) %>%
  arrange(state, city)

kable(cvg, row.names = TRUE, format = "latex", booktabs = TRUE, align = 'r') %>%
  kable_styling(latex_options = "scale_down")
```

total stops: `r prettyNum(sum(cvg$nrows), big.mark = ',', scientific=F)`

number of states represented: `r length(unique(cvg$state))`