### 2018-11-01 lib/coverage.Rmd
progress

---
title: "Open Policing Project: Coverage Rates"
classoption: landscape
---

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(kableExtra)
library(here)
source(here::here("lib", "coverage.R"))
cvg <- coverage()
ct <- filter(cvg, city != "Statewide")
kable(ct, row.names = TRUE, format = "latex", booktabs = TRUE, align = 'r') %>%
kable_styling(latex_options = "scale_down")
```
total city stops: `r prettyNum(sum(ct$nrows), big.mark = ',', scientific = F)`

number of states with city data: `r length(unique(ct$state))`

\newpage

```{r echo = FALSE, message = FALSE, warning = FALSE}
st <- filter(cvg, city == "Statewide")
kable(st, row.names = TRUE, format = "latex", booktabs = TRUE, align = 'r') %>%
kable_styling(latex_options = "scale_down")
```
total state stops: `r prettyNum(sum(st$nrows), big.mark = ',', scientific = F)`