### 2018-11-02 lib/prima_facie_stats.Rmd
progress

---
title: "Open Policing Project: Prima Facie Stats"
classoption: landscape
---

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(kableExtra)
library(here)
source(here::here("lib", "prima_facie_stats.R"))
stats <- prima_facie_stats()
tbl <- filter(stats, city != "Statewide")
kable(tbl, row.names = TRUE, format = "latex", booktabs = TRUE, align = 'r') %>%
kable_styling(latex_options = "scale_down")
```

\newpage

```{r echo = FALSE, message = FALSE, warning = FALSE}
tbl <- filter(stats, city == "Statewide")
kable(tbl, row.names = TRUE, format = "latex", booktabs = TRUE, align = 'r') %>%
kable_styling(latex_options = "scale_down")
```