### 2018-11-10 vodnotebook.Rmd
running vod regressions. before code clean up.

---
title: "Vod Report"
author: "Samantha Robertson"
date: "11/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source(here::here("lib", "opp.R"))
source(here::here("lib", "veil_of_darkness_test.R"))
```

```{r}
cities_w_data_problems <- c("Green Bay", "Camden", "Bakersfield", "Little Rock")
```

## Load data & clean up

```{r}
data <- 
  read_rds(here::here("cache", "aggregated_cities_wsunset.Rds")) %>% 
  filter(!(city %in% cities_w_data_problems)) %>% 
  filter(minute < sunset_minute | minute > sunset_minute + 30)

subgeos <- 
  data %>% 
  group_by(city) %>% 
  summarise_at(
    vars(
      neighborhood, 
      region, 
      precinct, 
      reporting_area, 
      district, 
      beat, 
      sector, 
      police_grid_number
    ), 
    n_distinct,
    na.rm = TRUE
  ) %>% 
  gather(-city, key = "subgeo", value = "n_distinct") %>% 
  filter(
    n_distinct >= 5, n_distinct <= 30, 
    (city != "Dallas" | subgeo != "district")
  ) %>% 
  select(-n_distinct)

get_subgeo_cities <- function(name) {
  subgeos %>% 
    filter(subgeo == name) %>% 
    pull(city)
}

region_cities <- get_subgeo_cities("region")
precinct_cities <- get_subgeo_cities("precinct")
district_cities <- get_subgeo_cities("district")
beat_cities <- get_subgeo_cities("beat")
sector_cities <- get_subgeo_cities("sector")

data <-
  data %>% 
  filter(city %in% pull(subgeos, city)) %>% 
  mutate(
    subgeography = case_when(
      city %in% region_cities ~ region,
      city %in% precinct_cities ~ precinct,
      city %in% district_cities ~ district,
      city %in% beat_cities ~ beat,
      city %in% sector_cities ~ sector
    )
  ) %>% 
  filter(!is.na(subgeography)) %>% 
  unite("geo_adj_var", city, subgeography, remove = FALSE)
```

### Find best times

```{r}
data %>% 
  ggplot(aes(time, fill = is_dark)) +
   geom_density(alpha = 1/2)
```

## Visualize

```{r}
plot_prop_minority_by_time(
  data, 
  title = "All stops", 
  min_time = hms::hms(hours = 18, min = 30), 
  max_time = hms::hms(hours = 19, min = 30)
)
```

```{r}
plot_prop_minority_by_time(
  data, 
  title = "All stops", 
  min_time = hms::hms(hours = 18, min = 30), 
  max_time = hms::hms(hours = 19, min = 30),
  smooth_method = "lm"
)
```

```{r}
data %>% filter(is_dst_period == TRUE) %>% count(city, sort = TRUE)
```

Use city with most stops around DST: Philadelphia

Get days of daylight savings:
```{r}
dst_dates <- 
  data %>% 
  filter(city == "Philadelphia", is_dst_period == TRUE) %>% 
  arrange(date) %>% 
  mutate(
    year = year(date),
    prev_dst = lead(is_dst),
    season = if_else(month < 6, "spring", "fall")
  ) %>% 
  filter(prev_dst != is_dst) %>% 
  select(year, dst_date = date, season)
```

Find time period  
```{r}
data %>% 
  group_by(date) %>% 
  summarise(sunset = min(sunset)) %>% 
  mutate(prev_sunset = lag(sunset)) %>% 
  filter(date %in% pull(dst_dates, dst_date)) %>% 
  mutate(season = if_else(month(date) < 6, "spring", "fall")) %>% 
  gather(sunset:prev_sunset, key = "day", value = "sunset_time") %>% 
  ggplot(aes(date, sunset_time, color = day, group = day)) +
  geom_point() +
  geom_line() +
  facet_grid(. ~ season)
```

5:15-5:30 in the fall and 6:10-6:25 in the spring

```{r}
data %>% 
  filter(city == "Philadelphia", is_dst_period == TRUE) %>% 
  mutate(
    year = year(date), 
    season = if_else(month > 6, "fall", "spring")
  ) %>% 
  left_join(dst_dates, by = c("year", "season")) %>% 
  mutate(days_since_dst = date - dst_date) %>%
  filter(
    days_since_dst < days(15),
    days_since_dst > days(-15),
    season == "spring" & time < hms::hms(hours = 18, min = 25) & time > hms::hms(hours = 18, min = 10) |
      season == "fall" & time < hms::hms(hours = 17, min = 30) & time > hms::hms(hours = 17, min = 15)
  ) %>% 
  group_by(days_since_dst, season, is_dark) %>% 
  summarise(prop_minority = sum(is_minority_demographic) / n(), n = n()) %>% 
  ggplot(aes(days_since_dst, prop_minority, color = is_dark))+
  geom_vline(xintercept = days(0)) +
  geom_point(aes(size = n), alpha = 1/2) +
  geom_smooth(se = FALSE, method = "lm") +
  facet_grid(. ~ season) +
  labs(
    title = "Philadelphia (2015-2018)",
    subtitle = "Time periods: 5:15-5:30 in the Fall & 6:10-6:25 in the Spring"
  )
```

Same for Nashville:
Get days of daylight savings:
```{r}
dst_dates_nashville <- 
  data %>% 
  filter(city == "Nashville", is_dst_period == TRUE) %>% 
  arrange(date) %>% 
  mutate(
    year = year(date),
    prev_dst = lead(is_dst),
    season = if_else(month < 6, "spring", "fall")
  ) %>% 
  filter(prev_dst != is_dst) %>% 
  select(year, dst_date = date, season)
```

```{r}
data %>% 
  group_by(date) %>% 
  summarise(sunset = min(sunset)) %>% 
  mutate(prev_sunset = lag(sunset)) %>% 
  filter(date %in% pull(dst_dates_nashville, dst_date)) %>% 
  mutate(season = if_else(month(date) < 6, "spring", "fall")) %>% 
  gather(sunset:prev_sunset, key = "day", value = "sunset_time") %>% 
  ggplot(aes(date, sunset_time, color = day, group = day)) +
  geom_point() +
  geom_line() +
  facet_grid(. ~ season)
```
Filter to 2011 - 2014: 5:30-5:45 fall, 6:10-6:25 spring
```{r}
data %>% 
  filter(city == "Nashville", is_dst_period == TRUE, year(date) <= 2014, year(date) >= 2011) %>% 
  mutate(
    year = year(date), 
    season = if_else(month > 6, "fall", "spring")
  ) %>% 
  left_join(dst_dates, by = c("year", "season")) %>% 
  mutate(days_since_dst = date - dst_date) %>%
  filter(
    days_since_dst < days(15),
    days_since_dst > days(-15),
    season == "spring" & time < hms::hms(hours = 18, min = 45) & time > hms::hms(hours = 18, min = 30) |
      season == "fall" & time < hms::hms(hours = 17, min = 45) & time > hms::hms(hours = 17, min = 30)
  ) %>% 
  group_by(days_since_dst, season, is_dark) %>% 
  summarise(prop_minority = sum(is_minority_demographic) / n(), n = n()) %>% 
  ggplot(aes(days_since_dst, prop_minority, color = is_dark))+
  geom_vline(xintercept = days(0)) +
  geom_point(aes(size = n), alpha = 1/2) +
  geom_smooth(se = FALSE, method = "lm") +
  facet_grid(. ~ season) +
  labs(
    title = "Nashville (2011-2014)",
    subtitle = "Time periods: 5:30-5:45 in the Fall; 6:30-6:45 in the Spring"
  )
```

## Regressions

### All data

```{r}
log(calculate_k(data))
```

```{r}
simple_model <-
  data %>% 
  group_by(is_dark, minute) %>% 
  summarise(
    n = n(),
    k = sum(is_minority_demographic)
  ) %>% 
  glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6), family = "binomial")

broom::tidy(simple_model)
```

```{r}
geo_model <-
  data %>% 
  group_by(is_dark, minute, geo_adj_var) %>% 
  summarise(
    n = n(), 
    k = sum(is_minority_demographic)
  ) %>% 
  glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6) + geo_adj_var, family = "binomial")

broom::tidy(geo_model)
```

### DST

```{r}
log(calculate_k(data %>% filter(is_dst_period == TRUE)))
```

```{r}
dst_simple_model <-
  data %>% 
  filter(is_dst_period == TRUE) %>% 
  group_by(is_dark, minute) %>% 
  summarise(
    n = n(),
    k = sum(is_minority_demographic)
  ) %>% 
  glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6), family = "binomial")

broom::tidy(dst_simple_model)
```

```{r}
dst_geo_model <-
  data %>% 
  filter(is_dst_period == TRUE) %>% 
  group_by(is_dark, minute, geo_adj_var) %>% 
  summarise(
    n = n(),
    k = sum(is_minority_demographic)
  ) %>% 
  glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6) + geo_adj_var, family = "binomial")

broom::tidy(dst_geo_model)
```

### Separately on each city

```{r}
reg <- function(city_name){
  model <-
    data %>% 
    filter(city == city_name) %>% 
    group_by(is_dark, minute, geo_adj_var) %>% 
    summarise(
      n = n(), 
      k = sum(is_minority_demographic)
    ) %>% 
    glm(formula = cbind(k, n-k) ~ is_dark + ns(minute, df = 6) + geo_adj_var, family = "binomial")

  broom::tidy(model) %>% 
    filter(term == "is_darkTRUE") %>% 
    mutate(city = city_name) %>% 
    select(city, estimate, std.error)
}

by_city <- 
  subgeos %>% 
  pull(city) %>% 
  map_dfr(~reg(.))

by_city %>% 
  filter(city != "Plano") %>% 
  mutate(low = estimate - std.error, high = estimate + std.error) %>% 
  ggplot(aes(reorder(city, estimate), estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = low, ymax = high))
```