### 2019-01-25 lib/bunching_report.Rmd

```{r, echo = FALSE}
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
source("bunching_test.R")

over_limit <- 40

display <- function(bt) {
  print(bt$results$plots$over)
  print(bt$results$plots$bunching)
  print(bt$results$plots$lenience)
  print(bt$results$plots$over_by_demographic_and_lenience)
  print(bt$results$plots$difference)
  print(bt$results$plots$coefficient)
}
```

# Decisions

1. Across all cities and states, we only had coverage in Oklahoma City, Dallas,
and Plano; and for these, we only had coverage for race and sex; we don't have
ages here
2. Non-lenient officers were defined to be those with $\le$ 2% of their
tickets at bunch points (same as original paper)
3. Used 10 MPH over as the only bunching point for these cities, since adding
15 MPH and 20 MPH over as bunching points didn't allow for differentiation
between lenient and non-lenient officers, i.e. very, very few officers had
fewer than 2% of their tickets at any of these 'bunch' points; it also seems
likely that the higher values are a mix of bunching and the thresholds at which
police really start stopping people; the original paper used 9 MPH over, since
10 MPH over corresponds to the first jump in the fee schedule; however, our
distributions are quite different from the original paper's, and not only in
this regard

\newpage

# Original FL data

Bunching point at 9MPH

```{r fl, fig.height=3.75, fig.width=7}
tbl <- haven::read_dta(here::here("data", "bunching", "Cite_05_15_Use.dta")) %>% 
  rename(
      subject_race = Race,
      subject_age = Age,
      officer_id = OfficerId, 
      speed = ActualSpeed, 
      posted_speed = PostedSpeed,
      vehicle_make = VehicleMake
    ) %>% 
  mutate(subject_race = plyr::mapvalues(
    subject_race, 
    from = c("B", "H", "W"), 
    to = c("black", "hispanic", "white")
  ))
tbl <-
  tbl %>%
  mutate(
    age_group = as.integer(subject_age / 10),
    hour_of_day = if_else(Hour == 99, NA_real_, Hour)
  )
bt <- bunching_test(
  tbl,
  CountyNum,
  hour_of_day,
  age_group,
  vehicle_make,
  over_limit = over_limit,
  bunching_points = c(9)
)
display(bt)
kable(bt$results$difference_in_difference)
```

\newpage

# Oklahoma City

Speeding Fine Schedule:

1-10 MPH over: \$172

11-19 MPH over: \$203

20+ MPH over: \$213

https://www.okc.gov/departments/municipal-court/fines-fees

```{r ok, fig.height=3.75, fig.width=7}
tbl <- opp_load_clean_data("ok", "oklahoma city")
tbl <-
  tbl %>%
  mutate(
    age_group = as.integer(subject_age / 10),
    hour_of_day = hour(hms(time))
  )
bt <- bunching_test(
  tbl,
  division,
  hour_of_day,
  age_group,
  vehicle_make,
  over_limit = over_limit
)
display(bt)
kable(bt$results$difference_in_difference)
```

\newpage

# Dallas

Speeding Fine Schedule:

1-10 MPH over: \$201

11-15 MPH over: \$226

16-20 MPH over: \$251

20 MPH over: \$304

https://dallascityhall.com/departments/courtdetentionservices/DCH%20Documents/OffenseCode_FinesFees_20180705.pdf

```{r dallas, fig.height=3.75, fig.width=7}
tbl <- opp_load_clean_data("tx", "dallas")
tbl <- mutate(tbl, hour_of_day = hour(hms(time)))
bt <- bunching_test(
  tbl,
  region,
  hour_of_day,
  vehicle_type,
  over_limit = over_limit
)
display(bt)
kable(bt$results$difference_in_difference)
```

\newpage

# Plano

Speeding Fine Schedule:

1-10 MPH over: \$204

11-20 MPH over: \$204 + \$10 * (MPH over - 10)

21-25 MPH over: \$304 + \$40 * (MPH over - 20)

https://www.plano.gov/364/Fine-Payment-Information

Used bunching value of 10 MPH over

```{r plano, fig.height=3.75, fig.width=7}
tbl <- opp_load_clean_data("tx", "plano")
tbl <- mutate(tbl, hour_of_day = hour(hms(time)))
# NOTE: location is null 99.97% of the time where speed is defined
bt <- bunching_test(
  tbl,
  hour_of_day,
  vehicle_make,
  over_limit = over_limit
)
display(bt)
kable(bt$results$difference_in_difference)
```