### 2020-02-28 lib/pfs_summary_table.Rmd
removing pfs summary table from output repot

---
title: PFS Summary
classoption: landscape
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
source(here::here("lib", "utils.R"))
source(here::here("lib", "format_plots_for_paper.R"))
library(tidyverse)
library(patchwork)
library(knitr)
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
pfs <- read_rds(here::here("results", "prima_facie_stats.rds"))
```

```{r search_rates_summary}
pfs$summary_table %>% 
  mutate(agency = if_else(City == "--", "state", "city")) %>% 
  group_by(agency) %>% 
  mutate(ct = 1:n()) %>% 
  ungroup() %>% 
  select(-agency) %>% 
  select(ct, everything()) %>%
  kable("latex")
```