### 2020-03-04 lib/paper_results.Rmd
Add aesthetic tweaks

---
title: A large-scale analysis of racial disparities in police stops across the United
  States
classoption: landscape
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
source(here::here("lib", "utils.R"))
source(here::here("lib", "format_plots_for_paper.R"))
library(tidyverse)
library(patchwork)
library(knitr)
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
pfs <- read_rds(here::here("results", "prima_facie_stats.rds"))
vod <- read_rds(here::here("results", "veil_of_darkness.rds"))
disp <- read_rds(here::here("results", "disparity.rds"))
mj <- read_rds(here::here("results", "marijuana_legalization_analysis.rds"))
mjf <- format_mj(mj)
```

\section{Basic Numbers}

We have collected and released a dataset of 
`r comma_num(pfs$counts$collected$n_stops)` traffic stops carried out by 
`r pfs$counts$collected$n_states` (`r comma_num(pfs$counts$collected$n_stops_states)`)
state patrol agencies and
`r pfs$counts$collected$n_cities` (`r comma_num(pfs$counts$collected$n_stops_cities)`)
municipal police departments.

We analyzed a unique dataset detailing 
`r comma_num(pfs$counts$analyzed$n_stops)` traffic stops carried out by 
`r pfs$counts$analyzed$n_states` (`r comma_num(pfs$counts$analyzed$n_stops_states)`)
state patrol agencies and
`r pfs$counts$analyzed$n_cities` (`r comma_num(pfs$counts$analyzed$n_stops_cities)`)
municipal police departments.

\section{Prima Facie Stats}

Stop rates:
```{r stop_rates}
pfs$rates$stop %>%
  mutate(
    agency = if_else(city == "Statewide", "state patrol", "municipal pd"),
    var = annual_stop_rate * (1 - annual_stop_rate) / population
  ) %>% 
  group_by(agency, subject_race) %>%
  summarize(
    stop_rate = mean(annual_stop_rate),
    std_error = sqrt(sum(var)) / n(),
    lower_ci = stop_rate - 1.96 * std_error,
    upper_ci = stop_rate + 1.96 * std_error
  ) %>%
  kable()
```

Search rates:
Number of state patrols with data for search analysis: 
`r pfs$rates$search %>% filter(city == "Statewide") %>% select(state) %>% n_distinct`

Number of city departments with data for search analysis: 
`r pfs$rates$search %>% filter(city != "Statewide") %>% select(state, city) %>% n_distinct`

```{r search_rates_cis}
pfs$rates$search %>% 
mutate(
  agency = if_else(city == "Statewide", "state patrol", "municipal pd"),
  var = search_rate * (1 - search_rate) / n
) %>% 
group_by(agency, subject_race) %>%
summarize(
  average_search_rate = mean(search_rate),
  std_error = sqrt(sum(var)) / n(),
  lower_ci = average_search_rate - 1.96 * std_error,
  upper_ci = average_search_rate + 1.96 * std_error
) %>%
ungroup() %>%
kable()
```

\newpage

\section{Summary Table}

total: `r pfs$summary_table$Stops %>% str_replace_all(",", "") %>% as.numeric() %>% sum %>% comma_num()`

\newpage

\section{Veil of Darkness}

Number of stops in 7:00-7:15pm panel:
```{r}
vod$full$plots$states$TX$`7:00pm`$data$n %>% sum() %>% comma_num()
```

Number of stops in 7:15-7:30pm panel:
```{r}
vod$full$plots$states$TX$`7:15pm`$data$n %>% sum() %>% comma_num()
```

Number of stops in 7:30-7:45pm panel:
```{r}
vod$full$plots$states$TX$`7:30pm`$data$n %>% sum() %>% comma_num()
```

Total number of stops in 3 panels:
```{r}
comma_num(vod$full$plots$states$TX$`7:00pm`$data$n %>% sum() + vod$full$plots$states$TX$`7:15pm`$data$n %>% sum() +  vod$full$plots$states$TX$`7:30pm`$data$n %>% sum())
```

```{r veil_of_darkness-plot-trio, fig.height=4, fig.width=9}
d <- rbind(
  vod$full$plots$states$TX$`7:00pm`$data,
  vod$full$plots$states$TX$`7:15pm`$data,
  vod$full$plots$states$TX$`7:30pm`$data
) %>% 
  mutate(time_window = str_c(
    time_str, " to ", minute_to_time(rounded_minute + 15)
  ))

d %>% 
  ggplot(aes(minutes_since_dark, proportion_minority)) +
  geom_point(aes(size = n)) +
  geom_smooth(
    aes(y = avg_p_minority, color = is_dark),
    method = "lm", se = F, linetype = "dashed"
  ) +
  geom_vline(xintercept = -25, linetype = "dotted") +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_ribbon(
    data = filter(d, is_dark),
    aes(ymin = avg_p_minority - se, ymax = avg_p_minority + se),
    alpha=0.3
  ) +
  geom_ribbon(
    data = filter(d, !is_dark),
    aes(ymin = avg_p_minority - se, ymax = avg_p_minority + se),
    alpha=0.3
  ) +
  scale_x_continuous(
    "Minutes since dusk",
    limits = c(-90, 60),
    breaks = seq(-90, 60, 30)
  ) +
  scale_y_continuous(
    str_c("Percent of stopped drivers who are black"),
    limits = c(0.15, 0.35),
    breaks = seq(0.0, 1.0, 0.05), 
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_color_manual(values = c("blue", "blue")) +
  theme_bw(base_size = 16) +
  theme(
    legend.position = 'none',
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  facet_grid(cols = vars(time_str))
```

\subsection{Main model results}

```{r veil_of_darkness-main_mod}
vod$dst %>% 
  select(-data, -base_controls) %>% 
  filter(spline_degree == 6, agency_control, !interact_time_loc) %>% 
  kable()
```

Point estimates: `r vod$dst %>% filter(spline_degree == 6, agency_control, !interact_time_loc) %>% pull(is_dark)`

95% CI (lower): `r vod$dst %>% filter(spline_degree == 6, agency_control, !interact_time_loc) %>% pull(is_dark) - 1.96*(vod$dst %>% filter(spline_degree == 6, agency_control, !interact_time_loc) %>% pull(std_error))`

95% CI (upper): `r vod$dst %>% filter(spline_degree == 6, agency_control, !interact_time_loc) %>% pull(is_dark) + 1.96*(vod$dst %>% filter(spline_degree == 6, agency_control, !interact_time_loc) %>% pull(std_error))`

\newpage
\subsection{Model results for the appendix/footnote reference}

Number of robustness check models with results not statistically significant: 
`r vod$dst %>% filter(is_dark + 1.96*std_error >= 0) %>% nrow`

```{r veil_of_darkness-appendix_mods}
vod$dst %>%
  select(-data, -base_controls) %>%
  mutate(
    l_ci = is_dark - 1.96 * std_error,
    u_ci = is_dark + 1.96 * std_error
  ) %>% 
  arrange(agency_control, interact_time_loc, spline_degree) %>% 
  select(agency, is_dark, std_error, l_ci, u_ci, p_value, everything()) %>%
  kable()
```

\newpage

\section{Search Disparity Analysis}

State patrol agencies analyzed: 
`r disp$state$outcome$results$hit_rates$geography %>% unique()`

Total number of state patrol searches: 
`r disp$state$outcome$results$hit_rates$n_search_conducted %>% sum() %>% comma_num()`

Municipal departments analyzed:
`r disp$city$outcome$results$hit_rates$geography %>% unique()`

Total number of municipal pd searches: 
`r disp$city$outcome$results$hit_rates$n_search_conducted %>% sum() %>% comma_num()`

\subsection{Outcome Test}

City and State outcomes:
```{r disparity-hit_rates-tables}
rbind(
disp$city$outcome$results$hit_rates %>% 
  group_by(geography, subject_race) %>% 
  summarize(
    hit_rate = weighted.mean(hit_rate, n_search_conducted),
    var = hit_rate * (1 - hit_rate) / sum(n_search_conducted)
  ) %>% 
  group_by(subject_race) %>% 
  summarize(
    geography = "cities",
    hit_rate = mean(hit_rate),
    std_error = sqrt(sum(var)) / n(),
    lower_ci = hit_rate - 1.96 * std_error,
    upper_ci = hit_rate + 1.96 * std_error
  ),
disp$state$outcome$results$hit_rates %>% 
  group_by(geography, subject_race) %>% 
  summarize(
    hit_rate = weighted.mean(hit_rate, n_search_conducted),
    var = hit_rate * (1 - hit_rate) / sum(n_search_conducted)
  ) %>% 
  group_by(subject_race) %>% 
  summarize(
    geography = "states",
    hit_rate = mean(hit_rate),
    std_error = sqrt(sum(var)) / n(),
    lower_ci = hit_rate - 1.96 * std_error,
    upper_ci = hit_rate + 1.96 * std_error
  )
) %>% 
select(geography, everything()) %>%
kable()
```
\newpage

```{r disparity-city-hit_rates-plot, fig.height=6, fig.width=10}
axis_min <- max(
  min(
    disp$city$outcome$plots$aggregate$data$minority_rate,
    disp$city$outcome$plots$aggregate$data$majority_rate
  ) - 0.05, 0
)
axis_max <- min(
  max(
    disp$city$outcome$plots$aggregate$data$minority_rate,
    disp$city$outcome$plots$aggregate$data$majority_rate
  ) + 0.05, 1
)
disp$city$outcome$plots$aggregate + 
  theme(legend.position = "none", plot.title = element_blank()) + 
  scale_x_continuous(
    limits = c(axis_min, axis_max),
    labels = scales::percent_format(accuracy = 1), expand = c(0,0)
  ) + 
  scale_y_continuous(
    limits = c(axis_min, axis_max),
    labels = scales::percent_format(accuracy = 1),expand = c(0,0)
  ) 
```

\newpage

```{r disparity-state-hit_rates-plot, fig.height=6, fig.width=10}
axis_min <- max(
  min(
    disp$state$outcome$plots$aggregate$data$minority_rate,
    disp$state$outcome$plots$aggregate$data$majority_rate
  ) - 0.03, 0
)
axis_max <- min(
  max(
    disp$state$outcome$plots$aggregate$data$minority_rate,
    disp$state$outcome$plots$aggregate$data$majority_rate
  ) + 0.03, 1
)

disp$state$outcome$plots$aggregate + 
  theme(
    legend.position = "none",
    plot.title = element_blank()
  ) +
  scale_x_continuous(
    limits = c(axis_min, axis_max),
    labels = scales::percent_format(accuracy = 1), expand = c(0,0)
  ) + 
  scale_y_continuous(
    limits = c(axis_min, axis_max),
    labels = scales::percent_format(accuracy = 1),expand = c(0,0)
  ) 
```

\newpage

\subsection{Threshold test results}

city then state

Scaling priors by 0.5x:
```{r disparity-thresholds-tables-0_5}
kable(disp$city$threshold$prior_scaling_factor_0.5$results$aggregate_thresholds)
kable(disp$state$threshold$prior_scaling_factor_0.5$results$aggregate_thresholds)
```

Priors at 1.0 (normal):
```{r disparity-thresholds-tables-1_0}
kable(disp$city$threshold$prior_scaling_factor_1$results$aggregate_thresholds)
kable(disp$state$threshold$prior_scaling_factor_1$results$aggregate_thresholds)
```

Scaling priors by 1.5x:
```{r disparity-thresholds-tables-1_5}
kable(disp$city$threshold$prior_scaling_factor_1.5$results$aggregate_thresholds)
kable(disp$state$threshold$prior_scaling_factor_1.5$results$aggregate_thresholds)
```

\newpage
Threshold plot for normal priors:

```{r disparity-city-thresholds-plot, fig.height=5, fig.width=10}
axis_min <- max(
  min(
    disp$city$threshold$prior_scaling_factor_1$plots$aggregate$data$minority_rate,
    disp$city$threshold$prior_scaling_factor_1$plots$aggregate$data$majority_rate
  ) - 0.03, 0
)
axis_max <- min(
  max(
    disp$city$threshold$prior_scaling_factor_1$plots$aggregate$data$minority_rate,
    disp$city$threshold$prior_scaling_factor_1$plots$aggregate$data$majority_rate
  ) + 0.03, 1
)
disp$city$threshold$prior_scaling_factor_1$plots$aggregate + 
  theme(legend.position = "none", plot.title = element_blank()) + 
  scale_x_continuous(
    limits = c(axis_min, axis_max), breaks = seq(0, .5, 0.05),
    labels = scales::percent_format(accuracy = 1), expand = c(0,0)
  ) + 
  scale_y_continuous(
    limits = c(axis_min, axis_max), breaks = seq(0, .5, 0.05),
    labels = scales::percent_format(accuracy = 1),expand = c(0,0)
  ) 
```

\newpage
Normal priors:

```{r disparity-state-thresholds-plot, fig.height=5, fig.width=10}
disp$state$threshold$prior_scaling_factor_1$plots$aggregate + 
  theme(legend.position = "none", plot.title = element_blank())
```

\newpage
Normal priors:

```{r disparity-city-ppc-search, fig.height=5, fig.width=8}
disp$city$threshold$prior_scaling_factor_1$ppc$search_rate + 
  scale_y_continuous(
    "Search rate prediction error\n",
    limits = c(-.016, .016), breaks = seq(-.5, .5, 0.005),
    labels = scales::percent_format(accuracy = 0.1)
) +
    scale_x_continuous(
        "\nPredicted search rate",
        limits = c(0, .27), breaks = seq(0, .5, 0.05),
        labels = scales::percent_format(accuracy = 1), expand = c(0,0)
    ) +
  theme(plot.title = element_blank())
```

\newpage
Normal priors:

```{r disparity-city-ppc-hit, fig.height=5, fig.width=8}
disp$city$threshold$prior_scaling_factor_1$ppc$hit_rate +
    scale_y_continuous(
        "Hit rate prediction error\n",
        limits = c(-.21, .21), breaks = seq(-.5, .5, 0.05),
        labels = scales::percent_format(accuracy = 1)
    ) +
    scale_x_continuous(
        "\nPredicted hit rate",
        limits = c(0, .37), breaks = seq(0, .5, 0.05),
        labels = scales::percent_format(accuracy = 1), expand = c(0,0)
    ) +
    theme(plot.title = element_blank())
```

\newpage
Normal priors:

```{r disparity-state-ppc-search, fig.height=5, fig.width=8}
disp$state$threshold$prior_scaling_factor_1$ppc$search_rate +
  scale_y_continuous(
        "Search rate prediction error\n",
        limits = c(-.021, .021), breaks = seq(-.5, .5, 0.005),
        labels = scales::percent_format(accuracy = .1)
    ) +
    scale_x_continuous(
        "\nPredicted search rate",
        limits = c(0, .11), breaks = seq(0, .5, 0.02),
        labels = scales::percent_format(accuracy = 1), expand = c(0,0)
    ) +
  theme(plot.title = element_blank())
```

\newpage
Normal priors:

```{r disparity-state-ppc-hit, fig.height=5, fig.width=8}
disp$state$threshold$prior_scaling_factor_1$ppc$hit_rate + 
  scale_y_continuous(
        "Hit rate prediction error\n",
        limits = c(-.30, .30), breaks = seq(-.5, .5, 0.05),
        labels = scales::percent_format(accuracy = 1)
    ) +
    scale_x_continuous(
        "\nPredicted hit rate",
        limits = c(0, .65), breaks = seq(0, .8, 0.1),
        labels = scales::percent_format(accuracy = 1), expand = c(0,0)
    ) +
  theme(plot.title = element_blank())
```

\newpage

\section{Marijuana Legalization Analysis}

```{r mj_table}
kable(mj$tables$search_rate_difference_in_difference_coefficients)
```

\newpage
```{r mj_search_rates_treatment, fig.height=3, fig.width=8}
sr <- mj$plots$search_rates
mjf$treatment_search_rates
```

CO: `r sr$CO$count`

WA: `r sr$WA$count`

decrease:
```{r mj_decrease}
mjf$treatment_discretionary_searches_with_no_contraband %>%
kable()
```

\newpage

```{r mj_misdemeanor_rates_treatment, fig.height=3, fig.width=8}
md <- mj$plots$misdemeanor_rates
mjf$treatment_misdemeanor_rates
```

CO: `r md$CO$count`

WA: `r md$WA$count`

\newpage

```{r mj_search_rates_control, fig.height=5.75, fig.width=10}
mjf$control_search_rates
```
AZ: `r sr$AZ$count`

CA: `r sr$CA$count`

FL: `r sr$FL$count`

MA: `r sr$MA$count`

MT: `r sr$MT$count`

NC: `r sr$NC$count`

OH: `r sr$OH$count`

RI: `r sr$RI$count`

SC: `r sr$SC$count`

TX: `r sr$TX$count`

VT: `r sr$VT$count`

WI: `r sr$WI$count`

\newpage

```{r mj_inferred_threshold_plot, fig.height=5, fig.width=10}
mj$plots$inferred_threshold_changes$prior_scaling_factor_1$plot +
  theme_bw(base_size = 20) +
  theme(
    panel.grid = element_blank(),
    legend.background=element_rect(fill="transparent"),
    legend.position=c(0.65, 0.25)
  )
```

\newpage

Prior robustness checks

Scaling factor 0.5:

WA rhat = `r mj$plots$inferred_threshold_changes$prior_scaling_factor_0.5$metadata$wa_rhat`

WA n_eff = `r mj$plots$inferred_threshold_changes$prior_scaling_factor_0.5$metadata$wa_n_eff`

CO rhat = `r mj$plots$inferred_threshold_changes$prior_scaling_factor_0.5$metadata$co_rhat`

CO n_eff = `r mj$plots$inferred_threshold_changes$prior_scaling_factor_0.5$metadata$co_n_eff`

```{r mj_inferred_threshold_plot_0_5_prior_scale, fig.height=5, fig.width=10}
mj$plots$inferred_threshold_changes$prior_scaling_factor_0.5$plot
```

Scaling factor 1.5:

WA rhat = `r mj$plots$inferred_threshold_changes$prior_scaling_factor_1.5$metadata$wa_rhat`

WA n_eff = `r mj$plots$inferred_threshold_changes$prior_scaling_factor_1.5$metadata$wa_n_eff`

CO rhat = `r mj$plots$inferred_threshold_changes$prior_scaling_factor_1.5$metadata$co_rhat`

CO n_eff = `r mj$plots$inferred_threshold_changes$prior_scaling_factor_1.5$metadata$co_n_eff`

```{r mj_inferred_threshold_plot_1_5_prior_scale, fig.height=5, fig.width=10}
mj$plots$inferred_threshold_changes$prior_scaling_factor_1.5$plot
```